% Analysis

# Packages
Load the packages needed to the analysis

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(deSolve)
library(ggthemes)
library(cowplot)
library(viridis)
```

# `ggplot2` global theme

```{r}
theme_set(theme_light())
```

# functions

The fuction for simulation of the measures  and particle filter are called here. 
```{r}
source("logistic.R")
source("particle_filter2.R")
```

# Estimation for logistic model

Here we set the initial values for simulation of each type of infection rate behavior.

```{r}
logi_setup = matrix(
  c("Constant", 0.2,
  "Increasing", 0.05,
  "Decreasing", 0.3,
  "Sinusoidal", 0.2,
  "Random", 0.2),
  nrow=5,ncol=,byrow = TRUE) 

```

Run the particle filter for each type of infection rate and for each time interval

```{r warning=FALSE}
noise = c(0.1, 0.25)

logistic_all3 = data.frame()
for(j in 1:2){
logistic_all2 = data.frame()
for(k in seq(1,10, by = 2)){

logistic_all = data.frame()
for( i in 1:5){
  set.seed(5)
data = logistic(N= 60, dt=0.5, y0=0.001, r=as.numeric(logi_setup[i,2]), sd = noise[j], inf = i)
data = data %>% 
  filter(time %in% c(seq(0,60, by = k)))
data_logi = data.frame(infection_type = as.factor(logi_setup[i,1]),
                   SIR_filter(model = 1,
                   guess_r = as.numeric(logi_setup[i,2]),
                   Nparti=100,
                   measures = data$Randon_intensity,
                   time = data$time,
                   sd_meas = 0.25,
                   sd_par = 0.15,
                   sd_model = 0.005),
                   y = data$Intensity,
                   inf_rate = data$inf_rate)  

logistic_all = logistic_all %>% 
  bind_rows(data_logi)
}

logistic_all = logistic_all %>% 
  mutate(time_interval = k)
  
logistic_all2 = logistic_all2 %>% 
  bind_rows(logistic_all) 
}
  logistic_all2 = logistic_all2 %>% 
    mutate(noise = noise[j])
  
 logistic_all3 =  logistic_all3 %>% 
   bind_rows(logistic_all2) 
   
}

logistic_all3 = logistic_all3 %>% 
  mutate(noise = case_when(noise == 0.10 ~ "\u03b1 =  0.10",
                           noise == 0.25 ~ "\u03b1 =  0.25")) %>% 
   mutate(time_interval2 = time_interval) %>% 
  mutate(time_interval = case_when(time_interval == 1 ~ "\u0394t =  1",
                                   time_interval == 3 ~ "\u0394t =  3",
                                   time_interval == 5 ~ "\u0394t =  5",
                                   time_interval == 7 ~ "\u0394t =  7",
                                   time_interval == 9 ~ "\u0394t =  9"))
```


## Visualization
```{r fig.height=10, fig.width=10}

logistic_all3 %>% 
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsiro), ymax = (lbdsiro), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,inf_rate, color = infection_type),
            size = 1.2)+
  geom_point(aes(time,sinti, shape = as.factor(noise)),
             size = 2, alpha = 0.5)+
  facet_grid(time_interval+noise~infection_type, scales = "free_y")+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",
       y = "Apparent infection rate")+
  theme(legend.position = "none",text = element_text(size = 14))+
  scale_y_continuous(breaks = seq(-2,2, 0.2))

ggsave("figs/logistic_air.png", dpi=300, height = 12, width = 8)
```

```{r fig.height=10, fig.width=10}
logistic_all3 %>% 
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsir), ymax = (lbdsir), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,y, color = infection_type),
            size = 1.2)+
  geom_point(aes(time,xestsir),shape=1,
              size = 1.5, alpha = 0.7)+
  facet_grid(time_interval+noise~infection_type)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",
       y = "Severity")+
  theme(legend.position = "none",text = element_text(size = 16))+
  scale_y_continuous(breaks = seq(0,1, 0.25))
ggsave("figs/logistic_curve.png", dpi=300, height = 12, width = 8)
```

## noised DPC

```{r}
logistic_all3 %>% 
  ggplot()+
  geom_line(aes(time,realSmeas, color = infection_type),
            size = 1.2)+

  facet_grid(time_interval+noise~infection_type)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",
       y = "Severity")+
  theme(legend.position = "none",text = element_text(size = 16))+
  scale_y_continuous(breaks = seq(0,1, 0.25))
ggsave("figs/logistic_noised.png", dpi=300, height = 12, width = 8)
```




## Acuracy

```{r warning=FALSE}
a = logistic_all3 %>% 
  group_by(infection_type,time_interval2, noise) %>%
  mutate(rmsi = (inf_rate - sinti)^2,
         maei = abs(inf_rate - sinti)) %>% 
  summarise(RMS = sqrt((1/(length(inf_rate)))*sum(rmsi, na.rm = T)),
            MAE = mean(maei)) %>% 
  mutate(model = "Logistic")

acuracy_logi = a


```



# Calculate the Apparent infection rate

## Simulated epidemics
```{r}
calc_r_log = logistic_all3 %>%
  group_by(infection_type,time_interval,noise) %>% 
  mutate(r_calc = (log(realSmeas/(1-realSmeas)) - log((lag(realSmeas,1)/(1-(lag(realSmeas,1))))))/(time - lag(time,1))) %>% 
  mutate(model = "Logistic")


calculated_r = calc_r_log
 
```


  Logistic
```{r fig.height=10, fig.width=10}

calculated_r %>%
  ggplot()+
  geom_line(aes(time,inf_rate, color = infection_type),
            size = 1.2)+
  geom_point(aes(time,r_calc, shape = as.factor(noise)),
             size = 2,
             alpha = 1)+
  facet_grid(time_interval+noise~infection_type, scales = "free_y")+
  scale_fill_viridis()+
  scale_color_colorblind()+
  labs(x = "Time",
       y = "Apparent infection rate")+
  guides(color= guide_legend("none"))+
  theme(text = element_text(size = 14), legend.position = "none")

 ggsave("figs/r_calc_logi.png", dpi=300, height = 12, width = 8)

```

## Acuracy

```{r warning=FALSE}
acuracy_calc = calculated_r%>% 
  filter(r_calc != is.na(r_calc)) %>% 
  group_by(infection_type, time_interval2,noise) %>% 
  mutate(rmsi = (inf_rate - r_calc)^2,
         maei = abs(inf_rate - r_calc)) %>% 
  summarise(RMS = sqrt((1/(length(inf_rate)))*sum(rmsi, na.rm = T)))
acuracy_calc
```

 bind data.frames of both models 
```{r}
acuracy_est = acuracy_logi

```
 
# Visualization 

## RMS 
 plotting the RMS 

```{r}
ggplot()+
  geom_line(data =acuracy_calc, aes(time_interval2, RMS, color = "Calculated", linetype = as.factor(noise)),
            size = 1.5)+
  geom_line(data =acuracy_est, aes(time_interval2, RMS, color = "Estimated", linetype = as.factor(noise)),
            size = 1.5)+
  scale_color_colorblind("AIR")+
  scale_x_continuous(breaks = seq(1,9,by=2))+
  scale_linetype_manual(values = c("solid","dashed"))+
  facet_grid(~infection_type)+
  guides(linetype=guide_legend(keywidth = 3, keyheight = 1),
         color=guide_legend(keywidth = 3, keyheight = 1))+
  labs(x = "Time interval", y = "RMSE", linetype  = "Noise")+
  theme(legend.position = "right",
        text = element_text(size = 14))

ggsave("figs/RMS.png", dpi=300, height = 5, width = 11)
```
## graph 1
```{r fig.height=8, fig.width=6, message=FALSE, warning=FALSE}
plot_grid(
  logistic_all3 %>%
  filter(time_interval == 1 & noise == 0.1 ) %>%
  filter(infection_type == c("Random")) %>%
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsiro), ymax = (lbdsiro), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,inf_rate),size = 1.5, color = "steelblue")+
  geom_point(aes(time,sinti),size = 2, alpha = 0.5)+
  # facet_grid(.~infection_type+noise)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",y = "AIR")+
  theme(legend.position = "none",text = element_text(size = 12),
        axis.title.x =element_text(color = "white"), axis.text.x = element_text(color = "white"))+
  scale_y_continuous(breaks = seq(-2,2, 0.2))+
  ylim(-0.8,1.4),
  
  calculated_r %>%
   filter(time_interval == 1 & noise == 0.1) %>%
  filter(infection_type == c("Random")) %>%
  ggplot()+
  geom_line(aes(time,inf_rate),size = 1.5, color = "steelblue")+
  geom_point(aes(time,r_calc),size = 2,alpha = 0.5)+
  # facet_grid(.~infection_type+noise)+
  scale_fill_viridis()+
  scale_color_colorblind()+
  labs(x = "Time",y = "AIR")+
  scale_y_continuous(breaks = seq(-2,2, 0.2),)+
  guides(color= guide_legend("none"))+
  ylim(-0.8,1.4)+
  theme(text = element_text(size = 12), legend.position = "none", axis.title.y =element_text(color = "white"), axis.text.y = element_text(color = "white"),
        axis.title.x =element_text(color = "white"), axis.text.x = element_text(color = "white")),
  
  logistic_all3 %>%
  filter(time_interval == 1 &  noise == 0.25 ) %>%
  filter(infection_type == c("Random")) %>%
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsiro), ymax = (lbdsiro), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,inf_rate),size = 1.5, color = "steelblue")+
  geom_point(aes(time,sinti),size = 2, alpha = 0.5)+
  # facet_grid(.~infection_type+noise)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",y = "AIR")+
  theme(legend.position = "none",text = element_text(size = 12),
        axis.title.x =element_text(color = "white"), axis.text.x = element_text(color = "white"))+
  scale_y_continuous(breaks = seq(-2,2, 0.2))+
  ylim(-0.8,1.4),
  
  calculated_r %>%
   filter(time_interval == 1 & noise == 0.25) %>%
  filter(infection_type == c("Random")) %>%
  ggplot()+
  geom_line(aes(time,inf_rate),size = 1.5, color = "steelblue")+
  geom_point(aes(time,r_calc),size = 2,alpha = 0.5)+
  # facet_grid(.~infection_type+noise)+
  scale_fill_viridis()+
  scale_color_colorblind()+
  labs(x = "Time",y = "AIR")+
  scale_y_continuous(breaks = seq(-2,2, 0.2),)+
  guides(color= guide_legend("none"))+
  ylim(-0.8,1.4)+
  theme(text = element_text(size = 12), legend.position = "none", axis.title.y =element_text(color = "white"),axis.text.y = element_text(color = "white"),
        axis.title.x =element_text(color = "white"), axis.text.x = element_text(color = "white")),
  
   logistic_all3 %>%
  filter(time_interval == 1 &  noise == 0.1 ) %>%
  filter(infection_type == c("Sinusoidal")) %>%
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsiro), ymax = (lbdsiro), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,inf_rate),size = 1.5, color = "steelblue")+
  geom_point(aes(time,sinti),size = 2, alpha = 0.5)+
  # facet_grid(.~infection_type+noise)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",y = "AIR")+
  theme(legend.position = "none",text = element_text(size = 12),
        axis.title.x =element_text(color = "white"), axis.text.x = element_text(color = "white"))+
  scale_y_continuous(breaks = seq(-2,2, 0.2))+
  ylim(-0.8,1.4),
  
  calculated_r %>%
   filter(time_interval == 1 & noise == 0.1) %>%
  filter(infection_type == c("Sinusoidal")) %>%
  ggplot()+
  geom_line(aes(time,inf_rate),size = 1.5, color = "steelblue")+
  geom_point(aes(time,r_calc),size = 2,alpha = 0.5)+
  # facet_grid(.~infection_type+noise)+
  scale_fill_viridis()+
  scale_color_colorblind()+
  labs(x = "Time",y = "AIR")+
  scale_y_continuous(breaks = seq(-2,2, 0.2),)+
  guides(color= guide_legend("none"))+
  ylim(-0.8,1.4)+
  theme(text = element_text(size = 12), legend.position = "none", axis.title.y =element_text(color = "white"),axis.text.y = element_text(color = "white"),
        axis.title.x =element_text(color = "white"), axis.text.x = element_text(color = "white")),
  
  logistic_all3 %>%
  filter(time_interval == 1 &  noise == 0.25 ) %>%
  filter(infection_type == c("Sinusoidal")) %>%
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsiro), ymax = (lbdsiro), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,inf_rate),size = 1.5, color = "steelblue")+
  geom_point(aes(time,sinti),size = 2, alpha = 0.5)+
  # facet_grid(.~infection_type+noise)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",y = "AIR")+
  theme(legend.position = "none",text = element_text(size = 12))+
  scale_y_continuous(breaks = seq(-2,2, 0.2))+
  ylim(-0.8,1.4),
  
  calculated_r %>%
   filter(time_interval == 1 & noise == 0.25) %>%
  filter(infection_type == c("Sinusoidal")) %>%
  ggplot()+
  geom_line(aes(time,inf_rate),size = 1.5, color = "steelblue")+
  geom_point(aes(time,r_calc),size = 2,alpha = 0.5)+
  # facet_grid(.~infection_type+noise)+
  scale_fill_viridis()+
  scale_color_colorblind()+
  labs(x = "Time",y = "AIR")+
  scale_y_continuous(breaks = seq(-2,2, 0.2),)+
  guides(color= guide_legend("none"))+
  ylim(-0.8,1.4)+
  theme(text = element_text(size = 12), legend.position = "none", axis.title.y =element_text(color = "white"),axis.text.y = element_text(color = "white")),
  
  ncol = 2,
  labels = "AUTO"

)
ggsave("figs/Fig3.png", dpi=300, height = 8, width = 6)
```
## graph 2
```{r fig.height=8, fig.width=6, message=FALSE, warning=FALSE}
plot_grid(
  logistic_all3 %>%
  filter(time_interval == 5 & noise == 0.1 ) %>%
  filter(infection_type == c("Random")) %>%
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsiro), ymax = (lbdsiro), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,inf_rate),size = 1.5, color = "steelblue")+
  geom_point(aes(time,sinti),size = 2, alpha = 0.5)+
  # facet_grid(.~infection_type+noise)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",y = "AIR")+
  theme(legend.position = "none",text = element_text(size = 12),
        axis.title.x =element_text(color = "white"), axis.text.x = element_text(color = "white"))+
  scale_y_continuous(breaks = seq(-2,2, 0.2))+
  ylim(-0.8,1.4),
  
  calculated_r %>%
   filter(time_interval == 5 & noise == 0.1) %>%
  filter(infection_type == c("Random")) %>%
  ggplot()+
  geom_line(aes(time,inf_rate),size = 1.5, color = "steelblue")+
  geom_point(aes(time,r_calc),size = 2,alpha = 0.5)+
  # facet_grid(.~infection_type+noise)+
  scale_fill_viridis()+
  scale_color_colorblind()+
  labs(x = "Time",y = "AIR")+
  scale_y_continuous(breaks = seq(-2,2, 0.2),)+
  guides(color= guide_legend("none"))+
  ylim(-0.8,1.4)+
  theme(text = element_text(size = 12), legend.position = "none", axis.title.y =element_text(color = "white"), axis.text.y = element_text(color = "white"),
        axis.title.x =element_text(color = "white"), axis.text.x = element_text(color = "white")),
  
  logistic_all3 %>%
  filter(time_interval == 5 &  noise == 0.25 ) %>%
  filter(infection_type == c("Random")) %>%
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsiro), ymax = (lbdsiro), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,inf_rate),size = 1.5, color = "steelblue")+
  geom_point(aes(time,sinti),size = 2, alpha = 0.5)+
  # facet_grid(.~infection_type+noise)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",y = "AIR")+
  theme(legend.position = "none",text = element_text(size = 12),
        axis.title.x =element_text(color = "white"), axis.text.x = element_text(color = "white"))+
  scale_y_continuous(breaks = seq(-2,2, 0.2))+
  ylim(-0.8,1.4),
  
  calculated_r %>%
   filter(time_interval == 5 & noise == 0.25) %>%
  filter(infection_type == c("Random")) %>%
  ggplot()+
  geom_line(aes(time,inf_rate),size = 1.5, color = "steelblue")+
  geom_point(aes(time,r_calc),size = 2,alpha = 0.5)+
  # facet_grid(.~infection_type+noise)+
  scale_fill_viridis()+
  scale_color_colorblind()+
  labs(x = "Time",y = "AIR")+
  scale_y_continuous(breaks = seq(-2,2, 0.2),)+
  guides(color= guide_legend("none"))+
  ylim(-0.8,1.4)+
  theme(text = element_text(size = 12), legend.position = "none", axis.title.y =element_text(color = "white"),axis.text.y = element_text(color = "white"),
        axis.title.x =element_text(color = "white"), axis.text.x = element_text(color = "white")),
  
   logistic_all3 %>%
  filter(time_interval == 5 &  noise == 0.1 ) %>%
  filter(infection_type == c("Sinusoidal")) %>%
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsiro), ymax = (lbdsiro), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,inf_rate),size = 1.5, color = "steelblue")+
  geom_point(aes(time,sinti),size = 2, alpha = 0.5)+
  # facet_grid(.~infection_type+noise)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",y = "AIR")+
  theme(legend.position = "none",text = element_text(size = 12),
        axis.title.x =element_text(color = "white"), axis.text.x = element_text(color = "white"))+
  scale_y_continuous(breaks = seq(-2,2, 0.2))+
  ylim(-0.8,1.4),
  
  calculated_r %>%
   filter(time_interval == 5 & noise == 0.1) %>%
  filter(infection_type == c("Sinusoidal")) %>%
  ggplot()+
  geom_line(aes(time,inf_rate),size = 1.5, color = "steelblue")+
  geom_point(aes(time,r_calc),size = 2,alpha = 0.5)+
  # facet_grid(.~infection_type+noise)+
  scale_fill_viridis()+
  scale_color_colorblind()+
  labs(x = "Time",y = "AIR")+
  scale_y_continuous(breaks = seq(-2,2, 0.2),)+
  guides(color= guide_legend("none"))+
  ylim(-0.8,1.4)+
  theme(text = element_text(size = 12), legend.position = "none", axis.title.y =element_text(color = "white"),axis.text.y = element_text(color = "white"),
        axis.title.x =element_text(color = "white"), axis.text.x = element_text(color = "white")),
  
  logistic_all3 %>%
  filter(time_interval == 5 &  noise == 0.25 ) %>%
  filter(infection_type == c("Sinusoidal")) %>%
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsiro), ymax = (lbdsiro), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,inf_rate),size = 1.5, color = "steelblue")+
  geom_point(aes(time,sinti),size = 2, alpha = 0.5)+
  # facet_grid(.~infection_type+noise)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",y = "AIR")+
  theme(legend.position = "none",text = element_text(size = 12))+
  scale_y_continuous(breaks = seq(-2,2, 0.2))+
  ylim(-0.8,1.4),
  
  calculated_r %>%
   filter(time_interval == 5 & noise == 0.25) %>%
  filter(infection_type == c("Sinusoidal")) %>%
  ggplot()+
  geom_line(aes(time,inf_rate),size = 1.5, color = "steelblue")+
  geom_point(aes(time,r_calc),size = 2,alpha = 0.5)+
  # facet_grid(.~infection_type+noise)+
  scale_fill_viridis()+
  scale_color_colorblind()+
  labs(x = "Time",y = "AIR")+
  scale_y_continuous(breaks = seq(-2,2, 0.2),)+
  guides(color= guide_legend("none"))+
  ylim(-0.8,1.4)+
  theme(text = element_text(size = 12), legend.position = "none", axis.title.y =element_text(color = "white"),axis.text.y = element_text(color = "white")),
  
  ncol = 2,
  labels = "AUTO"

)
ggsave("figs/Fig4.png", dpi=300, height = 8, width = 6)
```

