% Document analysis

# Packages
Load the packages needed to the analysis

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggthemes)
library(readxl)
library(cowplot)
library(viridis)
library(agricolae)
library(DescTools)
library(deSolve)
```

# `ggplot2` global theme

```{r}
theme_set(theme_light())
```

# functions

The fuction for simulation of the measures  and particle filter are called here. 
```{r}
source("logistic.R")
source("gompertz.R")
source("particle_filter2.R")
```

# Estimation for logistic model

Here we set the initial values for simulation of each type of infection rate behavior.

```{r}
logi_setup = matrix(
  c("Constant", 0.2,
  "Increasing", 0.05,
  "Decreasing", 0.3,
  "Sinusoidal", 0.2,
  "Random", 0.2),
  nrow=5,ncol=,byrow = TRUE) 

```

Run the particle filter for each type of infection rate and for each time interval

```{r warning=FALSE}
noise = c(0.1, 0.25)

logistic_all3 = data.frame()
for(j in 1:2){
logistic_all2 = data.frame()
for(k in seq(1,10, by = 2)){

logistic_all = data.frame()
for( i in 1:5){
  set.seed(5)
data = logistic(N= 60, dt=0.5, y0=0.001, r=as.numeric(logi_setup[i,2]), sd = noise[j], inf = i)
data = data %>% 
  filter(time %in% c(seq(0,60, by = k)))
data_logi = data.frame(infection_type = as.factor(logi_setup[i,1]),
                   SIR_filter(model = 1,
                   guess_r = as.numeric(logi_setup[i,2]),
                   Nparti=100,
                   measures = data$Randon_intensity,
                   time = data$time,
                   sd_meas = 0.25,
                   sd_par = 0.15,
                   sd_model = 0.005),
                   y = data$Randon_intensity,
                   inf_rate = data$inf_rate)  

logistic_all = logistic_all %>% 
  bind_rows(data_logi)
}

logistic_all = logistic_all %>% 
  mutate(time_interval = k)
  
logistic_all2 = logistic_all2 %>% 
  bind_rows(logistic_all) 
}
  logistic_all2 = logistic_all2 %>% 
    mutate(noise = noise[j])
  
 logistic_all3 =  logistic_all3 %>% 
   bind_rows(logistic_all2) 
   
}
```

```{r fig.height=10, fig.width=10}

logistic_all3 %>% 
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsiro), ymax = (lbdsiro), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,inf_rate, color = infection_type),
            size = 2)+
  geom_point(aes(time,sinti, shape = as.factor(noise)),
             size = 2)+
  facet_grid(time_interval+noise~infection_type)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",
       y = "Apparent infection rate")+
  theme(legend.position = "none")+
  scale_y_continuous(breaks = seq(-2,2, 0.2))

ggsave("figs/logistic_air.png", dpi=300, height = 10, width = 10)
```

```{r fig.height=10, fig.width=10}
logistic_all3 %>% 
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsir), ymax = (lbdsir), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,y, color = infection_type),
            size = 2)+
  geom_point(aes(time,xestsir, shape = as.factor(noise)),
             size = 2)+
  facet_grid(time_interval+noise~infection_type)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",
       y = "Intensity")+
  theme(legend.position = "none")+
  scale_y_continuous(breaks = seq(0,1, 0.2))
ggsave("figs/logistic_curve.png", dpi=300, height = 10, width = 10)
```


## Acuracy

```{r warning=FALSE}
a = logistic_all3 %>% 
  group_by(infection_type,time_interval, noise) %>%
  mutate(rmsi = (inf_rate - sinti)^2,
         maei = abs(inf_rate - sinti)) %>% 
  summarise(RMS = sqrt((1/(length(inf_rate)))*sum(rmsi, na.rm = T)),
            MAE = mean(maei)) %>% 
  mutate(model = "Logistic")

b = logistic_all3 %>% 
  group_by(infection_type,time_interval,noise) %>%
  filter(y < 1) %>% 
  mutate(rmsi = (inf_rate - sinti)^2,
         maei = abs(inf_rate - sinti)) %>% 
  summarise(RMS.1 = sqrt((1/(length(inf_rate)))*sum(rmsi, na.rm = T)),
            MAE.1 = mean(maei))
acuracy_logi = left_join(a, b)


```


# Estimation for Gompertz model

```{r}
gompi_setup = matrix(
  c("Constant", 0.1,
  "Increasing", 0.05,
  "Decreasing", 0.15,
  "Sinusoidal", 0.05,
  "Random", 0.2),
  nrow=5,ncol=,byrow = TRUE) 

```


```{r warning=FALSE}
noise = c(0.1, 0.25)

gompertz_all3 = data.frame()
for(j in 1:2){
gompertz_all2 = data.frame()
for(k in seq(1,10, by = 2)){
  
gompertz_all = data.frame()
for(i in 1:5){
 set.seed(5)
data = gompertz(N= 60, dt=0.5, y0=0.001,r=as.numeric(gompi_setup[i,2]), sd = noise[j], inf = i)
data = data %>% 
  filter(time %in% c(seq(1,60, by = k)))

data_gompi = data.frame(infection_type = as.factor(gompi_setup[i,1]),
                   SIR_filter(model = 2,
                   guess_r = as.numeric(gompi_setup[i,2]),
                   Nparti=100,
                   measures = data$Randon_intensity,
                   time = data$time,
                   sd_meas = 0.25,
                   sd_par = 0.15,
                   sd_model = 0.005),
                   y = data$Randon_intensity,
                   inf_rate = data$inf_rate)  

gompertz_all = gompertz_all %>% 
  bind_rows(data_gompi)

} 

gompertz_all = gompertz_all %>% 
  mutate(time_interval = k)
  
gompertz_all2 = gompertz_all2 %>% 
  bind_rows(gompertz_all) 
}  

  gompertz_all2 = gompertz_all2 %>% 
    mutate(noise = noise[j])
  
 gompertz_all3 =  gompertz_all3 %>% 
   bind_rows(gompertz_all2) 
   
}
```

```{r fig.height=10, fig.width=10}

gompertz_all3 %>% 
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsiro), ymax = (lbdsiro), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,inf_rate, color = infection_type),
            size = 2)+
  geom_point(aes(time,sinti, shape =  as.factor(noise)),
             size = 2)+
  facet_grid(time_interval+noise~infection_type)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",
       y = "Apparent infection rate")+
  theme(legend.position = "none",text = element_text(size = 14))+
  scale_y_continuous(breaks = seq(-2,2, 0.2))
 ggsave("figs/gompertz_air.png", dpi=300, height = 10, width = 10)
```
```{r fig.height=10, fig.width=10}
gompertz_all3 %>% 
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsir), ymax = (lbdsir), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,y, color = infection_type),
            size = 2)+
  geom_point(aes(time,xestsir, shape = as.factor(noise)),
             size = 2)+
  facet_grid(time_interval+noise~infection_type)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",
       y = "Intensity")+
  theme(legend.position = "none",text = element_text(size = 14))+
  scale_y_continuous(breaks = seq(-2,2, 0.2))
ggsave("figs/gompertz_curve.png", dpi=300, height = 10, width = 10)
```

## Acuracy

```{r warning=FALSE}
c = gompertz_all3 %>% 
  group_by(infection_type,time_interval, noise) %>%
  mutate(rmsi = (inf_rate - sinti)^2,
         maei = abs(inf_rate - sinti)) %>% 
  summarise(RMS = sqrt((1/(length(inf_rate)))*sum(rmsi, na.rm = T)),
            MAE = mean(maei)) %>% 
  mutate(model = "Gompertz")

d = gompertz_all3 %>% 
  group_by(infection_type,time_interval,noise) %>%
  filter(y < 1) %>% 
  mutate(rmsi = (inf_rate - sinti)^2,
         maei = abs(inf_rate - sinti)) %>% 
  summarise(RMS.1 = sqrt((1/(length(inf_rate)))*sum(rmsi, na.rm = T)),
            MAE.1 = mean(maei))  %>% 
  mutate(model = "Gompertz")

acuracy_gompi = left_join(c, d)
 
```


# Calculate the Apparent infection rate

## Simulated epidemics
```{r}
calc_r_log = logistic_all3 %>%
  group_by(infection_type,time_interval,noise) %>% 
  mutate(r_calc = (log(y/(1-y)) - log((lag(y,1)/(1-(lag(y,1))))))/(time - lag(time,1))) %>% 
  mutate(model = "Logistic")

calc_r_gomp =gompertz_all3 %>% 
  group_by(infection_type,time_interval,noise) %>% 
  mutate(r_calc = (-log(-log(y))--log(-log(lag(y,1))))/(time - lag(time,1))) %>%
  mutate(model = "Gompertz")

calculated_r = bind_rows(calc_r_log,calc_r_gomp)
calculated_r$model = factor(calculated_r$model, levels=c("Logistic","Gompertz"), labels=c("Logistic","Gompertz")) 
```



```{r fig.height=10, fig.width=10}

calculated_r %>%
  ggplot()+
  geom_line(aes(time,inf_rate, color = infection_type),
            size = 2)+
  geom_point(aes(time,r_calc),
             size = 2,
             alpha = 1)+
  facet_grid(time_interval+model~infection_type)+
  scale_fill_viridis()+
  scale_color_colorblind()+
  labs(x = "Time",
       y = "Apparent infection rate")+
  guides(color= guide_legend("none"))+
  xlim(0,60)+
  theme(text = element_text(size = 14))

  
# ggsave("figs/simulated_r_calc.png", dpi=300, height = 4, width = 6)

```
## Acuracy

```{r warning=FALSE}
acuracy_calc = calculated_r%>% 
  filter(r_calc != is.na(r_calc)) %>% 
  group_by(model,infection_type, time_interval,noise) %>% 
  mutate(rmsi = (inf_rate - r_calc)^2,
         maei = abs(inf_rate - r_calc)) %>% 
  summarise(RMS = sqrt((1/(length(inf_rate)))*sum(rmsi, na.rm = T)),
            CCC = CCC(inf_rate ,r_calc)$rho.c$est,
            MAE = mean(maei))
```

 bind data.frames of both models 
```{r}
acuracy_est =bind_rows(acuracy_logi,acuracy_gompi) 

```
 
# Visualization 

## RMS 
 plotting the RMS 

```{r}

ggplot()+
  geom_line(data =acuracy_calc, aes(time_interval, RMS, color = "Calculated", linetype = as.factor(noise)),
            size = 1.2)+
  geom_line(data =acuracy_est, aes(time_interval, RMS, color = "Estimated", linetype = as.factor(noise)),
            size = 1.2)+
  scale_color_colorblind("AIR")+
  scale_fill_colorblind()+
  facet_grid(model~infection_type)+
  labs(x = "Time interval", y = "RMS", linetype  = "Noise")+
  theme(legend.position = "bottom",text = element_text(size = 14))

ggsave("figs/RMS.png", dpi=300, height = 6, width = 10)
```


# Actual epidemics

Load the data 
```{r}
real = read_xlsx("curves_madden_book.xlsx")
```

## Estimated AIR
Run the Particle filter

```{r message=FALSE, warning=FALSE}
actual_all = data.frame()
real_matrix = as.matrix(real)
colnames(real_matrix) = NULL
for(i in 1:3){
  data_actual = data.frame(curve = as.factor(i),
                   SIR_filter(model = 1,
                   guess_r = 0.2,
                   Nparti=100,
                   measures = real_matrix[,i+1],
                   time = real$time,
                   sd_meas = 0.2,
                   sd_par = .5,
                   sd_model = 0.05,
                   dt= 7),
                   y = real_matrix[,i+1])  

actual_all = actual_all %>% 
  bind_rows(data_actual)
}
```
 Visualization
```{r}
plot_grid(  actual_all %>% 
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsir), ymax = (lbdsir), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,y, color = curve),
            size = 2)+
  geom_point(aes(time,xestsir),
             size = 2, 
             alpha = 0.3)+
  facet_wrap(~curve, nrow = 1)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",
       y = "Incidence")+
  theme(legend.position = "none",text = element_text(size = 14)),
  
actual_all %>% 
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsiro), ymax = (lbdsiro), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,sinti, color = curve),
             size = 2)+
  facet_wrap(~curve, nrow = 1)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",
       y = "Apparent infection rate")+
  theme(legend.position = "none",text = element_text(size = 14)),
nrow = 2)

# ggsave("figs/actual.png", dpi=300, height = 4, width = 6)
```



## Calculated AIR 


```{r}
plot_grid(  actual_all %>% 
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsir), ymax = (lbdsir), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,y),
            size = 1.2, color = "steelblue")+
  # geom_point(aes(time,xestsir),
  #            size = 2, 
  #            alpha = 1, color = "black")+
  facet_wrap(~curve, nrow = 1)+
  scale_fill_manual(values = "gray")+
  scale_color_colorblind()+
  labs(x = "Time",
       y = "Incidence")+
  theme(legend.position = "none"),

actual_all%>%  
  group_by(curve) %>% 
  mutate(r_calc = (log(y/(1-y)) - log((lag(y,1)/(1-(lag(y,1))))))/(time - lag(time,1))) %>% 
  ggplot()+
  geom_ribbon(aes(time, ymin = (ubdsiro), ymax = (lbdsiro), fill = "Ic 99%"), alpha = 0.5, stat = "identity") +
  geom_line(aes(time,sinti, linetype = "Estimated"),
             size = 1.2)+
  geom_point(aes(time,r_calc, shape = "Calculated"),
             size = 2,
             alpha  = 0.8)+
  facet_wrap(~curve, nrow = 1)+
  scale_fill_manual(values = "gray")+
  labs(x = "Time",fill = "", shape = "",linetype ="",
       y = "Apparent infection rate")+
  theme(legend.position = "bottom")+
  scale_y_continuous(breaks = seq(-2,2, 0.2)) ,
nrow = 2,
labels = c("A","B"), rel_heights  = c(0.7, 1)
)
 ggsave("figs/Actual_plots.png", dpi=300, height = 5, width = 8)
```


